---
title: "Focus on Ophelia (2019)"
author: "Ian D. Gow"
format: 
  html:
    toc: true
    code-fold: true
  pdf:
    colorlinks: true
    geometry:
      - left=2.5cm
      - right=2.5cm
    papersize: a4
    mainfont: TeX Gyre Pagella
    mathfont: TeX Gyre Pagella Math
    csl: journal-of-financial-economics.csl
---

```{r}
#| include: false
#| message: false
library(tidyverse)
library(duckdb)
library(conflicted)
library(dbplyr)
conflict_prefer("filter", "dplyr")
conflict_prefer("sql", "dbplyr")
```

```{r}
#| include: false
#| warning: false
db <- dbConnect(duckdb::duckdb())
movies <- tbl(db, "read_csv_auto('archive/rotten_tomatoes_movies.csv')")
reviews <- tbl(db, "read_csv_auto('archive/rotten_tomatoes_movie_reviews.csv')")

reviews_unique <-
  reviews |> 
  distinct() 

movies_unique <-
  movies |> 
  filter(!is.na(title)) |> 
  distinct() 
```

```{r}
#| echo: false
release_date <- 
  movies_unique |>
  filter(id == "ophelia_2019") |>
  select(releaseDateTheaters) |>
  pull()

ophelia_data <- 
  reviews_unique |> 
  filter(id == "ophelia_2019")
```

```{r}
#| eval: false
#| include: false
ophelia_data |> 
  window_order(creationDate) |> 
  mutate(is_fresh = as.integer(reviewState == "fresh")) |> 
  mutate(avg_rating = 100 * cummean(is_fresh), 
         num_ratings = cumsum(as.integer(!is.na(is_fresh)))) |> 
  filter(num_ratings >= 10) |>
  select(creationDate, avg_rating, num_ratings) |>
  pivot_longer(cols = -creationDate)
```

From [Vulture](https://www.vulture.com/article/rotten-tomatoes-movie-rating.html):

> Between October 2018 and January 2019, Rotten Tomatoes added eight reviews to Ophelia’s score. 
> Seven were favorable, and most came from critics who have reviewed at least one other Bunker 15 movie.
> The writer of a negative review says that Bunker 15 lobbied them to change it; if the critic wanted to "give it a (barely) overall positive then I do know the editors at Rotten Tomatoes and can get it switched," a Bunker 15 employee wrote. 
> I also discovered another negative review of Ophelia from this period that was not counted by Rotten Tomatoes, by a writer whose positive reviews of other Bunker 15 films have been recorded by the aggregator. 
> Ophelia climbed the Tomatometer to 62 percent, flipping from rotten to "fresh." 
The next month, the distributor IFC Films announced that it had acquired Ophelia for release in the U.S.

Some of these phenomena can be seen in @fig-reviews-time

```{r}
#| echo: false
ophelia_dates <-
  ophelia_data |> 
  summarize(min_date = min(creationDate, na.rm = TRUE),
            max_date = max(creationDate, na.rm = TRUE)) |>
    mutate(dates = generate_series(as.POSIXct(min_date), 
                                   as.POSIXct(max_date), 
                                   sql("'1 day'::interval"))) |>
  mutate(creationDate = as.Date(unnest(dates))) |>
  select(creationDate)
  
plot_data <-
  ophelia_dates |>
  left_join(ophelia_data, by = "creationDate") |> 
  window_order(creationDate) |> 
  mutate(is_fresh = as.integer(reviewState == "fresh")) |> 
  mutate(avg_rating = 100 * cummean(is_fresh), 
         num_ratings = cumsum(as.integer(!is.na(is_fresh)))) |> 
  filter(num_ratings >= 10) |>
  group_by(creationDate) |>
  filter(num_ratings == max(num_ratings, na.rm = TRUE)) |>
  ungroup() 
```

```{r}
#| echo: false
#| eval: true 
#| fig-height: 4
#| label: fig-reviews-time
#| layout-nrow: 2
#| fig-cap: "Ophelia on Rotten Tomatoes over time"
#| fig-subcap: 
#|   - Cumulative number of reviews
#|   - Tomato Meter
plot_data |> 
  ggplot(aes(x = creationDate)) + 
  geom_line(aes(y = num_ratings)) +
  geom_vline(xintercept = release_date, color = "red") +
  scale_x_date(date_minor_breaks = "1 month")
 
plot_data|> 
  ggplot(aes(x = creationDate)) + 
  geom_line(aes(y = avg_rating), color = "blue") +
  geom_vline(xintercept = release_date, color = "red") +
  geom_hline(yintercept = 59.5, color = "black") +
  scale_x_date(date_minor_breaks = "1 month")
```
```{r}
#| echo: false
#| eval: true 
#| fig-height: 4
#| label: fig-reviews-zoom
#| layout-nrow: 2
#| fig-cap: "Ophelia on Rotten Tomatoes around release"
#| fig-subcap: 
#|   - Cumulative number of reviews
#|   - Tomato Meter
plot_data |> 
  filter(between(creationDate, "2019-06-01", "2019-07-31")) |>
  ggplot(aes(x = creationDate)) + 
  geom_line(aes(y = num_ratings)) +
  geom_vline(xintercept = release_date, color = "red") +
  scale_x_date(date_minor_breaks = "1 day")
 
plot_data |> 
  filter(between(creationDate, "2019-06-01", "2019-07-31")) |>
  ggplot(aes(x = creationDate)) + 
  geom_line(aes(y = avg_rating), color = "blue") +
  geom_vline(xintercept = release_date, color = "red") +
  geom_hline(yintercept = 59.5, color = "black") +
  scale_x_date(date_minor_breaks = "1 day")
```




