---
title: "Quick look at Rotten Tomatoes data"
author: "Ian D. Gow"
format: 
  pdf:
    toc: true
  html:
    toc: true
    code-fold: true
---

Here I assume that you have saved the data available [here](https://www.kaggle.com/datasets/andrezaza/clapper-massive-rotten-tomatoes-movies-and-reviews) in a `archive` folder under this project.

I will use the `tidyverse` package for analysis.
I will use `duckdb` here because it makes queries very fast with very little overhead.^[Use `install.packages(c("tidyverse", "duckdb"))` if necessary.]

```{r}
#| message: false
library(tidyverse)
library(duckdb)
```

## Load data

Read the data into the database.

```{r}
db <- dbConnect(duckdb::duckdb())
movies <- tbl(db, "read_csv_auto('archive/rotten_tomatoes_movies.csv')")
reviews <- tbl(db, "read_csv_auto('archive/rotten_tomatoes_movie_reviews.csv')")
```

```{r}
movies |> count(id) |> count(n)
```

The result above tells us that are duplicate rows in the `movies` table.
If we focus on those with a value in `title`, it seems there are complete duplicates (i.e., all rows are duplicated).
So we can make a new table with unique entries.

```{r}
movies_unique <-
  movies |> 
  filter(!is.na(title)) |> 
  distinct() 
```

Now we have no duplicates.

```{r}
movies_unique |> 
  count(id) |> 
  filter(n > 1) |>
  count() |>
  pull()
```

A quick look at the data.

```{r}
movies_unique |>
  filter(!is.na(tomatoMeter)) |>
  count(tomatoMeter) |> 
  arrange(tomatoMeter) |>
  filter(tomatoMeter > 55) 
```

There are also some duplicates in `reviews`.
Let's make `reviews_unique`:

```{r}
reviews_unique <-
  reviews |> 
  distinct() 

reviews_unique |>
  count(reviewId) |> 
  filter(n > 1) |>
  inner_join(reviews, by = "reviewId") |>
  count() |>
  pull()
```

What values does `reviewState` take?
Are there missing values?

```{r}
reviews_unique |>
  count(reviewState) |>
  collect()

reviews_unique |>
  filter(is.na(reviewState)) |>
  count() |>
  pull()
```

## Calculating Tomato Score

Let's calculate tomato score.

```{r}
scores_calc <-
  reviews_unique |>
  mutate(pos_review = as.integer(reviewState == "fresh")) |>
  group_by(id) |>
  summarize(tomatoMeter = 100 * mean(pos_review, na.rm = TRUE),
            num_reviews = n()) |>
  compute()
```

Now let's compare calculated values with reported values for a sample.

```{r}
sample <- 
  tibble(id = c("american_werewolf_in_paris",
                "imperium",
                "welcome_to_happiness",
                "blood_glacier")) |>
  copy_to(db, df = _, name = "sample", overwrite = TRUE)
```

```{r}
sample |>
  inner_join(movies_unique, by = "id") |>
  select(id, title, tomatoMeter) |>
  inner_join(scores_calc, by = "id")
```

We can produce a histogram of Tomato Meter ratings using the scores in `movies`.

```{r}
#| label: fig-hist
#| fig-cap: Histogram of Tomato Meter ratings
movies_unique |>
  filter(!is.na(tomatoMeter)) |>
  ggplot(aes(x = tomatoMeter)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 100, 5))
```

We can also produce a histogram of Tomato Meter ratings using the scores we calculated ourselves.

```{r}
#| label: fig-calc-hist
#| fig-cap: Histogram of Tomato Meter ratings (calculated)
scores_calc |>
  ggplot(aes(x = tomatoMeter)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 100, 5))
```

```{r}
#| label: fig-calc-hist-alt
#| fig-cap: Histogram of Tomato Meter ratings (calculated, sub-sample)
movies_unique |>
  filter(!is.na(tomatoMeter)) |>
  select(-tomatoMeter) |>
  inner_join(scores_calc, by = "id") |>
  ggplot(aes(x = tomatoMeter)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 100, 5))
```

@fig-hist looks very different from @fig-calc-hist.
I'd guess that some criteria (e.g., minimum number of reviews) are imposed before including a Tomato Meter rating.
@fig-calc-hist-alt uses the calculated ratings, but limited to the subsample with `tomatoMeter` scores in `movies`.
It looks much more similar (though perhaps with too many 0 and 100 scores).
